On The Impact of Atoms of Confusion in JavaScript Code

### 2021-01-29 5_Threats.tex
revisao inicial da secao de ameacas

Since our research was conducted online with unknown participants, we had no way of confirming their levels of education and experience. Also, as we briefly discussed in Section 4.3, we did not have a way to prevent respondents from cheating, such as running the code on an interpreter, or consulting other people. This presents a threat to the validity of the histograms presented in Figures \ref{fig:degree} and \ref{fig:degree}.

Another threat to our results lies in the fact that some constructs might not be frequent, or present at all, in JavaScript code bases. The Automatic Semicolon Insertion and the Comma Operator atoms were not found in the code samples we consulted when formulating the questions, and we had to create blocks of code containing the atoms. The work done in Chapter 5 could serve as a starting point for detecting these atoms in a large data set of JavaScript files.

Regarding hypotheses H.2 \todo[inline]{Linkar as Hipoteses} and research question RQ.2, a potential problem with our method is that there was no strong incentive for subjects to think thoroughly about the questions. We observed a lack of engagement when we ran the survey with undergraduate students. Although our final subjects were voluntarily partaking in the survey, we could not be sure that, after some time taking the survey, respondents would become tired and stop thinking clearly about the code. As questions were assigned randomly, and contained different levels of difficulty, comparing a subject's answer with their other questions would not provide a reliable metric.

In one of the atoms, namely the Omitted Curly Braces, we intentionally removed indentation from the code, which is highly unusual, given that many programmers use automatic formatting in their code editors. This can introduce some level of artificiality to this atom's question. On the other hand, while describing JavaScript's grammar to develop the parser we used for the work in Chapter 5, we discovered that the language allows the programmer to omit the curly braces after \textit{if} statements, and insert multiple statements in the following line. This fact itself might constitute a source of confusion, which we leave to analyse in our future endeavors.

>>>>

Since our research was conducted online with unknown participants, and thus we had no way of confirming their
levels of education and experience. Also,  we did not have a way to prevent respondents from cheating,
such as running the code on an interpreter, or consulting other people. This presents a threat to the validity to the
conclusions of our work, even though we leverage some means to try to mitigate this threat (as discussed
in Section~\ref{method}).

Another threat to our results lies in the fact that some constructs
might not be frequent, or present at all, in JavaScript code bases.
The atom candidates Arithmetic as Logic and
the Comma Operator were rarely found in the
open source JavaScript code we mined in our
third study.

A potential problem with our method is that
there little incentive for subjects to think
thoroughly about the questions. We observed a lack of
engagement when we ran the survey with undergraduate students during
a pilot study. Although our final subjects were voluntarily
partaking in the survey, we could not be sure that, after some
time taking the survey, respondents would become tired and stop thinking
clearly about the code.

Finally, in one of the atoms, namely the Omitted Curly Braces,
we intentionally removed indentation from the original code,
which is highly unusual, given that many programmers use automatic
formatting in their code editors. This can introduce some level
of artificiality to this atom's question. Nonetheless,
we discovered that JavaScript does allow the programmer
to omit the curly braces after \textit{if} statements,
and insert multiple statements in the following line.
This fact itself might constitute
a source of confusion, which we leave to analyse
in our future endeavors.
